Galactic Chemical Evolution Models Favor an Extended Type Ia Supernova Delay-Time Distribution

### 2024-06-21 src/tex/ms.tex
[showyourwork] overleaf sync

We restrict our sample to red giant branch and red clump stars with high-quality spectra. Table \ref{tab:sample} lists our selection criteria, which largely follow from \citet{Hayden2015-ChemicalCartography}. This produces a final sample of \samplesize stars with {\bf calibrated} [O/Fe] and [Fe/H] abundance measurements. 
{\bf APOGEE stars were cross-matched with the {\it Gaia} Early Data Release 3 (EDR3) catalog \citep{Gaia2016-Mission,Gaia2021-EDR3} by the APOGEE team using Two Micron All Sky Survey \citep[2MASS;][]{Skrutskie2006-2MASS} identifiers, which were cross-matched with EDR3 by the {\it Gaia} collaboration. We use the \citet{Bailer-Jones2021-GaiaDistances} photo-geometric distance estimates}
% APOGEE stars are matched with their Bailer-Jones photo-geometric distance estimate from \textit{Gaia} Early Data Release 3 \citep{Gaia2016-Mission,Gaia2021-EDR3}, which we use 
to calculate galactocentric radius $R_{\rm gal}$ and midplane distance $z$, {\bf assuming a Sun--Galactic center distance $R_\odot=8.122$ kpc and height of the Sun above the midplane $z_\odot=20.8$ pc. Table \ref{tab:apogee-regions} lists the number of APOGEE stars in bins of $R_{\rm gal}$ and $|z|$. For Galactic regions with $R_{\rm gal}<5$ kpc or $R_{\rm gal}>13$ kpc, the median distance error exceeds 1 kpc but remains within our bin width of 2 kpc, and the vast majority of stars have much smaller distance uncertainties.}

@@

{\bf The origin of the Milky Way's \aFe bimodality is not currently understood. \citet{Sharma2021-RadialMigration} and \citet{Chen2023-RadialMixingRedux} showed that a GCE model with a continuous star formation history and radial migration can reproduce the observed abundance distributions. Similarly, \citet{Kubryk2015-RadialMigrationEvolution} and \citet{Prantzos2023-ThinThickDisks} argued that radial migration, combined with different SFHs for the inner and outer disk, can explain the \aFe bimodality. On the other hand, many studies have argued that the bimodality is instead driven by multiple bursts of gas infall \citep[e.g.,][]{Chiappini1997-TwoInfall,Mackereth2018-AlphaBimodality,Spitoni2019-TwoInfall,Spitoni2020-TwoInfall,Spitoni2021-TwoInfall}. We found that radial migration does not predict the \aFe bimodality on its own, in agreement with \citetalias{Johnson2021-Migration}.}

@@

% Software acknowledgements if not a AAS journal
% This work made use of Astropy:\footnote{http://www.astropy.org} a community-developed core Python package and an ecosystem of tools and resources for astronomy \citep{astropy:2013, astropy:2018, astropy:2022}.

>>>>

We restrict our sample to red giant branch and red clump stars with high-quality spectra. Table \ref{tab:sample} lists our selection criteria, which largely follow from \citet{Hayden2015-ChemicalCartography}. This produces a final sample of \samplesize stars with {\bf calibrated} [O/Fe] and [Fe/H] abundance measurements. 
{\bf APOGEE stars were cross-matched with the {\it Gaia} Early Data Release 3 (EDR3) catalog \citep{Gaia2016-Mission,Gaia2021-EDR3} by the APOGEE team using Two Micron All Sky Survey \citep[2MASS;][]{Skrutskie2006-2MASS} identifiers, which were cross-matched with EDR3 by the {\it Gaia} collaboration. We use the \citet{Bailer-Jones2021-GaiaDistances} photo-geometric distance estimates}
% APOGEE stars are matched with their Bailer-Jones photo-geometric distance estimate from \textit{Gaia} Early Data Release 3 \citep{Gaia2016-Mission,Gaia2021-EDR3}, which we use 
to calculate galactocentric radius $R_{\rm gal}$ and midplane distance $z$, {\bf assuming a Sun--Galactic center distance $R_\odot=8.122$ kpc and height of the Sun above the midplane $z_\odot=20.8$ pc. Table \ref{tab:apogee-regions} lists the number of APOGEE stars in bins of $R_{\rm gal}$ and $|z|$. For some Galactic regions with $R_{\rm gal}<5$ kpc or $R_{\rm gal}>13$ kpc, the median distance error exceeds 1 kpc but remains within our bin width of 2 kpc, and the vast majority of stars have much smaller distance uncertainties.}

@@

{\bf The origin of the Milky Way's \aFe bimodality remains disputed. Some authors have argued that the combination of inside-out growth and radial migration sufficiently explains the observed distribution \citep[e.g.,][]{Kubryk2015-RadialMigrationEvolution,Sharma2021-RadialMigration,Chen2023-RadialMixingRedux,Prantzos2023-ThinThickDisks}, while others have argued that multiple episodes of gas infall at early times are required \citep[e.g.,][]{Chiappini1997-TwoInfall,Mackereth2018-AlphaBimodality,Spitoni2019-TwoInfall,Spitoni2020-TwoInfall,Spitoni2021-TwoInfall}. In agreement with \citetalias{Johnson2021-Migration}, we find that a smooth SFH combined with radial migration does not suffice. We find that these parameter choices predict too many stars between the high- and low-$\alpha$ sequences, resulting in a broad but unimodal \aFe distribution at fixed [Fe/H].
% \citet{Sharma2021-RadialMigration} and \citet{Chen2023-RadialMixingRedux} showed that a GCE model with a continuous star formation history and radial migration can reproduce the observed abundance distributions. Similarly, \citet{Kubryk2015-RadialMigrationEvolution} and \citet{Prantzos2023-ThinThickDisks} argued that radial migration, combined with different SFHs for the inner and outer disk, can explain the \aFe bimodality. On the other hand, many studies have argued that the bimodality is instead driven by multiple bursts of gas infall \citep[e.g.,][]{Chiappini1997-TwoInfall,Mackereth2018-AlphaBimodality,Spitoni2019-TwoInfall,Spitoni2020-TwoInfall,Spitoni2021-TwoInfall}. We found that radial migration does not predict the \aFe bimodality on its own, in agreement with \citetalias{Johnson2021-Migration}.
}

@@
